---
title: 'AMS 597 Project'
author: "Michael Tartamella"
date: "5/5/2023"
output:
  word_document: default
---

Reads in our R package which we have created.
```{r}
library(biomMetaTartamella)
```

Simulates data to test our functions.
```{r}
set.seed(123)
p <- 100
data1 <- data.frame(group = sample(1:3, 200, replace = TRUE),
matrix(rnorm(p * 200), ncol = p))
data2 <- data.frame(group = sample(1:2, 150, replace = TRUE),
matrix(rnorm(p * 150), ncol = p))
data3 <- data.frame(group = sample(1:4, 300, replace = TRUE),
matrix(sample(0:2, size = 300 * p, replace = TRUE),
ncol = p))
colnames(data1) <- c("group", paste("gene", 1:p, sep = ""))
colnames(data2) = colnames(data3) <- colnames(data1)
input.data <- list(Data1 = data1, Data2 = data2, Data3 = data3)
```

Displays number of observations of each group across the three data sets.
```{r}
table(input.data$Data1$group)
table(input.data$Data2$group)
table(input.data$Data3$group)
```
First, we use three different functions to generate our p-values from each of our three different data sets.  We store these values into three different variables.  Next, I combine these p-values into one variable for ease of access, along their respective weights into one variable.  Then, I preallocate space to hold p-values for each one of our pooled p-value methods.  Next, I run a for loop to iterate through all our biomarkers using every single function created to pool our p-values and store I store these into their respective variable.  Finally, I test my "adjust" function on every one of our pooled p-values from each of the tests.
```{r}
pval1 <- pvals.manyGroups.anova.kw(data1)
pval2 <- pvals.twoGroups.t.wilcoxon(data2)
pval3 <- pvals.categorical.fisher.chisquared(data3)

combined.pvals <- cbind(pval1$p.value, pval2$p.value, pval3$p.value)
combined.n <- cbind(pval1$n, pval2$n, pval3$n)

fisher.pvalues <- numeric(nrow(combined.pvals))
stouffer.pvalues <- numeric(nrow(combined.pvals))
weighted.stouffer.pvalues <- numeric(nrow(combined.pvals))
lancaster.pvalues <- numeric(nrow(combined.pvals))
minp.pvalues <- numeric(nrow(combined.pvals))
maxp.pvalues <- numeric(nrow(combined.pvals))
weighted.fisher.pvalues <- numeric(nrow(combined.pvals))
for (i in 1:nrow(combined.pvals))
{
  fisher.pvalues[i] <- fisher.pool(combined.pvals[i, ])
  stouffer.pvalues[i] <- stouffer.pool(combined.pvals[i, ])
  weighted.stouffer.pvalues[i] <- weighted.stouffer.pool(combined.pvals[i, ], combined.n)
  lancaster.pvalues[i] <- lancaster.pool(combined.pvals[i, ], combined.n)
  minp.pvalues[i] <- minp.pool(combined.pvals[i, ])
  maxp.pvalues[i] <- maxp.pool(combined.pvals[i, ])
  weighted.fisher.pvalues[i] <- weighted.fisher.pool(combined.pvals[i, ])
}

adjusted.fisher.pvalues <- adjust(fisher.pvalues)
adjusted.stouffer.pvalues <- adjust(stouffer.pvalues)
adjusted.weighted.stouffer.pvalues <- adjust(weighted.stouffer.pvalues)
adjusted.lancaster.pvalues <- adjust(lancaster.pvalues)
adjusted.min.p.pvalues <- adjust(minp.pvalues)
adjusted.max.p.pvalues <- adjust(maxp.pvalues)
adjusted.weighted.fisher.pvalues <- adjust(weighted.fisher.pvalues)
```
